Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-11-23 prop_letter.tex
prop_letter.tex Updated by Prof. Ortiz...

We also investigate the second key aspect of imprecise quantum
measurement, which is its statistical nature as reflected in the
application of the Born rule determining the probability of a given
measurement outcome~\citep{Born1983bibTeX,Mermin2007,Jaeger2007}.
Here, Gleason's theorem establishes that there is no alternative to
the Born rule by demonstrating, using reasonable continuity arguments
in Hilbert spaces of dimension $d\ge3$, the existence of a unique
state $\rho$ consistent with the statistical predictions computed from
the Born rule~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Our question then concerns what happens when infinite precision cannot
be achieved.  We are able to show that, while a QIVPM incorporating
the effects of finite precision might not be consistent with Gleason's
unique state~$\rho$ on all projectors defined on a Hilbert
space~$\Hilb$ of dimension $d\ge3$, there is a mathematically precise
sense in which one recovers the original Gleason theorem
asymptotically.  Specifically, it is possible to construct a class of
QIVPMs representing bounded resources that is parameterized by the
size of the intervals. We then demonstrate that all QIVPMs in this
class are consistent with a non-empty ``ball'' of quantum states whose
distance are all within a distance proportional to the size of the
intervals, and as the size of the intervals characterizing the
uncertainties goes to zero, this ball of quantum states converges to a
point representing the unique state consistent with the Born rule. 
%% We furthermore establish a mathematical correlation between the sizes
%% of the intervals and the volume of states that are all consistent with
%% an imprecise measurement.  \andy{``volume of states" are not defined
%%   in the Gleason's section.}

@@

To summarize, a conventional quantum probability measure has an
uncountable range $[0,1]$. A QIVPM constructed by blurring such a
conventional quantum probability measure must also have an uncountable
range of intervals. Of course, any particular QIVPM, or any particular
experiment, will use a fixed collection of intervals appropriate for
the resources and precision of the particular experiment. 
\gerardo{We need to explain physically what it means?}
% The essential reason for getting an uncountable number of intervals is
% the convexity condition we imposed on QIVPMs. This condition is widely
% used in the classical setting but it is possible that it should be
% modified or weakened in the quantum case. If we take this approach and 
% % Even if $\ultramodular$ has to be ultramodular to guarantee
% % $\ultramodular\circ\mu$ a QIVPM, the composited function
% % $\ultramodular\circ\mu$ may still be a QIVPM when $\ultramodular$ is
% % not ultramodular. In this case, the composited function
% % $\ultramodular\circ\mu$ could map to non-trivial finite number of
% % intervals. For example, we can 
% weaken the convexity condition, Eq.~(\ref{eq:iota-Inclusion}), we can
% recover QIVPMs that only require countable and even finite collections
% of intervals. Formally, we weaken the convexity condition to the
% following additive condition: for any numbers $x_{0}$ and
% $x_{1}\in\left[0,1\right]$, their sum $x_{0}+x_{1}\in\left[0,1\right]$
% implies
% \begin{equation}
% \ultramodular\left(x_{0}+x_{1}\right)\subseteq\ultramodular\left(x_{0}\right)+\ultramodular\left(x_{1}\right)\,.\label{eq:iota-add}
% \end{equation}
% Then, given a three-dimensional Hilbert space, if $\ultramodular$
% satisfies Eqs.~(\ref{eq:iota-constraints}) and
% Eq.~(\ref{eq:iota-add}), then $\ultramodular\circ\mu$ is a
% QIVPM~\citep{TaiThesis2018}.  For example, consider an orthonormal
% basis $\{\ket{0},\ket{1},\ket{2}\}$, two states
% $\rho_{0}=\frac{1}{3}\proj{0}+\frac{1}{3}\proj{1}+\frac{1}{3}\proj{2}$
% and
% $\rho_{1}=\frac{7}{24}\proj{0}+\frac{1}{3}\proj{1}+\frac{3}{8}\proj{2}$,
% and a set of intervals:
% \begin{equation}
% \begin{aligned}\mathscr{I}_{2}={} & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\\
%  & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{3}{4},1\right],\necess\right\} \,.
% \end{aligned}
% \end{equation}
% Since the interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$
% defined in Table~\ref{tab:I_2} satisfies Eqs.~(\ref{eq:iota-constraints})
% and Eq.~(\ref{eq:iota-add})~\citep{TaiThesis2018}, $\ultramodular_{2}\circ\muB_{\rho_{0}}$
% and $\ultramodular_{2}\circ\muB_{\rho_{1}}$ are both
% QIVPMs.
% \begin{table}
% \caption{\label{tab:I_2}A interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$.}

>>>>

We also investigate the second key aspect of imprecise quantum
measurement, which is its statistical nature as reflected in the
application of the Born rule determining the probability of a given
measurement outcome~\citep{Born1983bibTeX,Mermin2007,Jaeger2007}.
Here, Gleason's theorem establishes that there is no alternative to
the Born rule by demonstrating, using reasonable continuity arguments
in Hilbert spaces of dimension $d\ge3$, the existence of a unique
state $\rho$ consistent with the statistical predictions computed from
the Born rule~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Our question then concerns what happens when infinite precision cannot
be achieved.  We are able to show that, while a QIVPM incorporating
the effects of finite precision might not be consistent with Gleason's
unique state~$\rho$ on all projectors defined on a Hilbert
space~$\Hilb$ of dimension $d\ge3$, there is a mathematically precise
sense in which one recovers the original Gleason theorem
asymptotically.  Specifically, it is possible to construct a class of
QIVPMs representing bounded resources that is parameterized by the
size of the intervals. We then demonstrate that all QIVPMs in this
class are consistent with a non-empty ``ball'' of quantum states whose
radius is defined by the maximal length of the
intervals characterizing the uncertainties. As the size of the intervals  
goes to zero, this ball of quantum states converges to a
point representing the unique state consistent with the Born rule and 
Gleason's theorem. 
%% We furthermore establish a mathematical correlation between the sizes
%% of the intervals and the volume of states that are all consistent with
%% an imprecise measurement.  \andy{``volume of states" are not defined
%%   in the Gleason's section.}

@@

To summarize, a conventional quantum probability measure has an
uncountable range $[0,1]$. A QIVPM constructed by blurring such a
conventional quantum probability measure must also have an uncountable
range of intervals. Of course, any particular QIVPM, or any particular
experiment, will use a fixed collection of intervals appropriate for
the resources and precision of the particular experiment. 
%\gerardo{We need to explain physically what it means?}
% The essential reason for getting an uncountable number of intervals is
% the convexity condition we imposed on QIVPMs. This condition is widely
% used in the classical setting but it is possible that it should be
% modified or weakened in the quantum case. If we take this approach and 
% % Even if $\ultramodular$ has to be ultramodular to guarantee
% % $\ultramodular\circ\mu$ a QIVPM, the composited function
% % $\ultramodular\circ\mu$ may still be a QIVPM when $\ultramodular$ is
% % not ultramodular. In this case, the composited function
% % $\ultramodular\circ\mu$ could map to non-trivial finite number of
% % intervals. For example, we can 
% weaken the convexity condition, Eq.~(\ref{eq:iota-Inclusion}), we can
% recover QIVPMs that only require countable and even finite collections
% of intervals. Formally, we weaken the convexity condition to the
% following additive condition: for any numbers $x_{0}$ and
% $x_{1}\in\left[0,1\right]$, their sum $x_{0}+x_{1}\in\left[0,1\right]$
% implies
% \begin{equation}
% \ultramodular\left(x_{0}+x_{1}\right)\subseteq\ultramodular\left(x_{0}\right)+\ultramodular\left(x_{1}\right)\,.\label{eq:iota-add}
% \end{equation}
% Then, given a three-dimensional Hilbert space, if $\ultramodular$
% satisfies Eqs.~(\ref{eq:iota-constraints}) and
% Eq.~(\ref{eq:iota-add}), then $\ultramodular\circ\mu$ is a
% QIVPM~\citep{TaiThesis2018}.  For example, consider an orthonormal
% basis $\{\ket{0},\ket{1},\ket{2}\}$, two states
% $\rho_{0}=\frac{1}{3}\proj{0}+\frac{1}{3}\proj{1}+\frac{1}{3}\proj{2}$
% and
% $\rho_{1}=\frac{7}{24}\proj{0}+\frac{1}{3}\proj{1}+\frac{3}{8}\proj{2}$,
% and a set of intervals:
% \begin{equation}
% \begin{aligned}\mathscr{I}_{2}={} & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\\
%  & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{3}{4},1\right],\necess\right\} \,.
% \end{aligned}
% \end{equation}
% Since the interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$
% defined in Table~\ref{tab:I_2} satisfies Eqs.~(\ref{eq:iota-constraints})
% and Eq.~(\ref{eq:iota-add})~\citep{TaiThesis2018}, $\ultramodular_{2}\circ\muB_{\rho_{0}}$
% and $\ultramodular_{2}\circ\muB_{\rho_{1}}$ are both
% QIVPMs.
% \begin{table}
% \caption{\label{tab:I_2}A interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$.}
